---
title: "Data Science Statistical Inference"
author: "Ben Kooi"
date: "20-8-2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
## Synopsis
This report describes the analysis of the TootGrowth-dataset.
The questions that should be answered are:  
1. Provide a basic summary of the data;   
2. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose;   
3. Conclusions and assumptions.
    
The most important conclusion of analyzing the ToothGrowth-dataset is that we can reject the null-hypothesis for len ~ dose, but for len ~ supp we can accept the null-hypothesis. In other words, dose matters by tootgrowth and supp not.
 
## Setting the environment

Load the needed packages.
```{r message=FALSE}
packages <- c("data.table","dplyr", "ggplot2")
sapply(packages, require, character.only=TRUE, quietly=TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 2 - Analyzing Toothgrowth Dataset

The Toothgrowth Dataset contains the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, OJ=orange juice or VC=ascorbic acid.

First we do a few basic analysis of the data.
```{r }
#change type of variable dose
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

#inspect the summary and the structure of the dataset
summary(ToothGrowth)
str(ToothGrowth)

# make boxplot to analyze the len ~ dose
p<-ggplot(ToothGrowth, aes(x=dose, y=len, fill=dose)) +
  geom_boxplot() +
  ggtitle("Toothgrowth per dose") +
  labs(title="Plot of sampled distribution", x ="Dose", y = "Lenght")
p
```

If we analyze the plot, it seems that len is strong dependant of dose. For proving this we perform a nopaired t.test of every combination of doses.

```{r }
# testing for signficance
t.test(len ~ dose, paired = FALSE, data = subset(ToothGrowth,(ToothGrowth$dose %in% c(0.5, 1))))
t.test(len ~ dose, paired = FALSE, data = subset(ToothGrowth,(ToothGrowth$dose %in% c(0.5, 2))))
t.test(len ~ dose, paired = FALSE, data = subset(ToothGrowth,(ToothGrowth$dose %in% c(1, 2))))
```
Every t.test has a outcome of a p-value<0.05. So we can reject the null-hypothesis. There is a significant difference in toothgrowth if we change the dose of vitamin C.

The next analysis is to determine if the type of supplement (supp) has a significant difference in toothgrowth. First we make a boxplot.

```{r}
#calculate mean and variance of both supplements
mvc<-mean(subset(ToothGrowth,ToothGrowth$supp=="VC")$len)
vvc<-var(subset(ToothGrowth,ToothGrowth$supp=="VC")$len)
moj<-mean(subset(ToothGrowth,ToothGrowth$supp=="OJ")$len)
voj<-var(subset(ToothGrowth,ToothGrowth$supp=="OJ")$len)

# make a boxplot len~supp
plot<-ggplot(ToothGrowth, aes(x=supp, y=len, fill=supp)) +
  geom_boxplot() +
  ggtitle("Toothgrowth per supplement") +
  labs(title="Plot of sampled distribution", x ="Supplement", y = "Lenght")
plot
```

The mean of the groups are different, so are the variance. The means are **`r mvc`** vs **`r moj`** and the variances are **`r vvc`** vs **`r voj`**

```{r}
# testing for signficance
t.test(len ~ supp, paired = FALSE, data = ToothGrowth, var.equal = FALSE)
```

If we perform a nopaired t.test of len ~ supp the p-value is close to 0.5, but greater. So we have to reject the null-hypothesis. So it seems to be that the type of supplement doesn't matter for toothgrowth.


